The journey of Multiple Sequence Alignment

### 2021-08-06 MSA_Notum_journey.Rmd
add sponge sequences and make a new WIF tree

# hmmbuild
```{bash}
hmmbuild --amino --cpu 4  my_data/NOTUM_try my_data/hmmsearch_env_coord_msa_muscle.fa 
hmmbuild --amino --cpu 4  my_data/WIF_try my_data/WIF_hmmsearch_env_coord_msa_muscle.fa 
hmmbuild --amino --cpu 4  my_data/wnt_try my_data/wnt_hmmsearch_env_coord_msa_muscle.fa 
```

@@

```
## Import hmmscan results into R
```{r}
second_line_fields <- c("target_name", "accession_1", "tlen", "query_name",
                        "accession_2", "qlen",   "E-value_full_seq", "score_full_seq",
                        "bias_full_seq",
                        "this_domain_#",  "of_this_domain",  "c-Evalue_this_domain",
                        "i-Evalue_this_domain",  "score_this_domain", "bias_this_domain", 
                        "from_hmm_coord", "to_hmm_coord", "from_ali_coord", "to_ali_coord", 
                        "from_env_coord", "to_env_coord", "acc", "description_of_target_1",
                        "description_of_target_2","description_of_target_3")

@@

>>>>

@@

## Import hmmscan results into R
```{r}
second_line_fields <- c("target_name", "accession_1", "tlen", "query_name",
                        "accession_2", "qlen",   "E-value_full_seq", "score_full_seq",
                        "bias_full_seq",
                        "this_domain_#",  "of_this_domain",  "c-Evalue_this_domain",
                        "i-Evalue_this_domain",  "score_this_domain", "bias_this_domain", 
                        "from_hmm_coord", "to_hmm_coord", "from_ali_coord", "to_ali_coord", 
                        "from_env_coord", "to_env_coord", "acc", "description_of_target_1",
                        "description_of_target_2","description_of_target_3")

@@

# Remake the WIF tree
We will use some sequences from sponges with the previous sequences

## import the new sequences
```{r}
seq_sponges <- readAAStringSet("WIF_sponge.fasta")
```

WIF_hmmsearch <- data.table::fread(cmd = "grep -v '^#' hmm_search/WIF_domtbl.out | awk -v OFS='\t'  '{for(i=1;i<=23;i++) printf $i\"\t\"; print \"\"}'",  header = FALSE, sep = "\t",  fill=TRUE)

names(WIF_hmmsearch)[1:23] <- second_line_fields
```

## Get only the filtered sequences 
```{r}
# filter for the hmmsearch env_coord sequences
fasta_db_fil_with_len <- fasta_db[names_fst_db %in% WIF_hmmsearch$target_name]

WIF_hmmsearch <- WIF_hmmsearch %>% 
  filter(!duplicated(target_name)) %>% 
  select(target_name, from_env_coord, to_env_coord) %>% 
  arrange(target_name)

identical(fasta_db_fil_with_len %>% names %>% str_remove(" .+"),
          WIF_hmmsearch$target_name)

fasta_db_fil_with_len <- pmap(list(fasta_db_fil_with_len %>% as.list, 
          WIF_hmmsearch %>% select(target_name, from_env_coord) %>% deframe %>% as.list(),
          WIF_hmmsearch %>% select(target_name, to_env_coord) %>% deframe %>% as.list()),
     subseq) %>% Biostrings::AAStringSet()

# cat the two files
fasta_db_fil_with_len_sponges <- c(fasta_db_fil_with_len,seq_sponges) %>% 
  writeXStringSet("WIF_and_sponges_sequences.fasta")
```

## perform Multiple Sequence Alignment with MUSCLE hmmsearch
```{r}
## env_coord
hmmsearch_env_coord_muscle <- msaMuscle(fasta_db_fil_with_len_sponges, type = "protein")
hmmsearch_env_coord_bio2mds <- msa::msaConvert(hmmsearch_env_coord_muscle, type = "bios2mds::align")
export.fasta(hmmsearch_env_coord_bio2mds, outfile = "WIFtree_withsponges_hmmsearch_env_coord_msa_muscle.fa", ncol = 60, open = "w")
```

iqtree -s my_data/WIFtree_withsponges_hmmsearch_env_coord_msa_muscle.fa -nt 4 -mset LG -pre my_data/my_hmmsearch_tree/WIFtree_withsponges_hmmsearch_env_coord_tree
```

## search which domains these sponges sequences have
```{bash}
hmmscan --cut_ga --domtblout my_data/hmm_search/Sponges_domains_domtbl.out my_data/pfam/Pfam-A.hmm my_data/WIF_sponge.fasta > /dev/null
```

```{bash}
hmmscan --cut_ga --domtblout my_data/hmm_search/WIF_Sponges_domains_domtbl.out my_data/pfam/Pfam-A.hmm my_data/WIF_and_sponges_sequences.fasta > /dev/null
```

## Import the hmmscan with the domains
```{r}
second_line_fields <- c("target_name", "accession_1", "tlen", "query_name",
                        "accession_2", "qlen",   "E-value_full_seq", "score_full_seq",
                        "bias_full_seq",
                        "this_domain_#",  "of_this_domain",  "c-Evalue_this_domain",
                        "i-Evalue_this_domain",  "score_this_domain", "bias_this_domain", 
                        "from_hmm_coord", "to_hmm_coord", "from_ali_coord", "to_ali_coord", 
                        "from_env_coord", "to_env_coord", "acc", "description_of_target_1",
                        "description_of_target_2","description_of_target_3")

WIF_sponges_hmmscan <- data.table::fread(cmd = "grep -v '^#' hmm_search/WIF_Sponges_domains_domtbl.out | awk -v OFS='\t'  '{for(i=1;i<=29;i++) printf $i\"\t\"; print \"\"}'",  header = FALSE, sep = "\t",  fill=TRUE)

names(WIF_sponges_hmmscan)[1:25] <- second_line_fields

WIF_sponges_hmmscan %>% 
  distinct(target_name, query_name) %>% 
  pivot_wider(names_from = target_name, values_from = target_name,values_fill = "0") %>% 
  unite(col = "domains",2:4) %>% 
  mutate(domains = domains %>% str_remove_all("0_") %>% str_remove_all("_0"),
         query_name = query_name %>% str_replace("\\|","_")) %>% 
  mutate(query_name_1 = query_name) %>% 
  unite(query_name, 1:2, sep = "__DOMAIN__") %>% view()
```
